---
title: "The factors for a good quality coffee!"
author: "Group 12"
#output: html_document
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA) 
```

```{r libraries, echo=FALSE}
library(tidyverse)
library(moderndive)
library(gapminder)
library(sjPlot)
library(stats)
library(jtools)
library(MASS)
library(knitr)
library(ggplot2)
library(dplyr)
library(janitor)
library(broom)
```

## Introduction {#sec:Intro}
The data set here, is coming from the Coffee Quality Database which is a database that contains information from the Coffee
Quality Institute which is a non-profit organisation working internationally to improve the quality of coffee and the
lives of the people who produce it.
Here, in this project, we are trying to answer the question "What influence do different features of coffee have on whether the quality of a batch of coffee is classified as good or poor?"  

For this purpose, firstly, we read the data set.

```{r reading data,echo=TRUE}
coffee_data <- read.csv("dataset12.csv")
```


## Data distribution {#sec:EDA}

#### The Country of origin of the coffee

We take a look at the percentages of good and bad qualities of coffees with respect to it's country of origin.
```{r check distribution origin}
#Check the percentage of good and poor quality coffee for each country of origin

coffee_data %>% 
  tabyl(country_of_origin, Qualityclass) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  adorn_ns()

```

In order to get a clear picture of which countries coffee beans are mostly of good and bad quality, we can plot the quality based on the country of origin in gradient scale.

```{r plotting origin}

#Plot the distribution of coffee quality class based on the origin

ggplot(coffee_data, aes(y = Qualityclass)) +
  geom_bar(aes(fill = country_of_origin), position = position_stack(reverse = TRUE)) +
  theme(legend.position = "top")+
  scale_color_gradient(low = "yellow", high = "green")+
  labs(x = "Proportion of Country of Origin", y = "Quality of Coffee")
```

From the figure above, we can see that the color gradient is larger in the red area in case of the good quality coffee while that of blue is dominant in case of bad quality coffee.But the difference is not highly significant.








###  The Aroma of the coffee

Let's consider the next variable of interest, which is the aroma of the coffee. Here we want to see how the aroma of the coffee is distributed around the quality of the coffee.


```{r plotting aroma1}

#Plot the distribution of coffee quality class based on the aroma
#First plot without rescaling
ggplot(data = coffee_data, aes(x = Qualityclass, y = aroma)) +
  geom_boxplot() +
  labs(title = "Quality score of coffee samples by aroma",
       x = "Aroma",
       y = "Qualityclass")
```



From the boxplot above, it is hard to get the picture of the distribution, because of the presence of outliers in this distribution. So we try to rescale the boxplot to get better visual of the distribution.



```{r plotting aroma2}

#Rescaling the plot
ggplot(data = coffee_data, aes(x = Qualityclass, y = aroma)) +
  geom_boxplot() +
  scale_y_continuous(limits = quantile(coffee_data$aroma, c(0.05, 0.95)))+
  labs(title = "Quality score of coffee samples by aroma (Rescale)",
       x = "Quality Class",
       y = "Aroma")
```

From the boxplot above, we can see that a low value of aroma of the coffee corresponds to poor quality.

Please note that for the rest of boxplot figure, we use the rescaled data in order to make the clearer representation of the distribution.

### Flavor

Here, we consider the next variable, flavor of the coffee with respect to the quality.
```{r plotting flavor}
ggplot(data = coffee_data, aes(x = Qualityclass, y = flavor)) +
  geom_boxplot() +
  scale_y_continuous(limits = quantile(coffee_data$flavor, c(0.05, 0.95)))+
  labs(title = "Quality score of coffee samples by flavor (Rescale)",
       x = "Quality Class",
       y = "flavor")
```

From the boxplot above, we can understand that a coffee is labelled as good if it has a high flavor grade.


### Acidity

Considering acidity grade with respect to the class of quality.

```{r plotting acidity}
ggplot(data = coffee_data, aes(x = Qualityclass, y = acidity)) +
  geom_boxplot() +
  scale_y_continuous(limits = quantile(coffee_data$acidity, c(0.05, 0.95)))+
  labs(title = "Quality score of coffee samples by acidity (Rescale)",
       x = "Quality Class",
       y = "acidity")
```

We can see that higher the acidity grade, better the coffee quality.


### Category 2 defects

This variable contains the number of times category 2 type defect was found in the batch of coffee beans tested.
```{r plotting category two defect}
ggplot(data = coffee_data, aes(x = Qualityclass, y =    category_two_defects)) +
  geom_boxplot() +
  scale_y_continuous(limits =   quantile(coffee_data$category_two_defects, c(0.05, 0.95)))+
  labs(title = "Quality score of coffee samples by category_two_defects (Rescale)",
       x = "Quality Class",
       y = "category_two_defects")
```

As we can see, category 2 type defects were more in poor quality coffee beans.And
there is no difference mean from the two distribution. For the futher analysis, we
consider to remove this variable.


### Mean altitude of the growers farm



```{r plotting altitude mean (meters)}
ggplot(data = coffee_data, aes(x = Qualityclass, y =    altitude_mean_meters)) +
  geom_boxplot() +
  scale_y_continuous(limits =   quantile(coffee_data$altitude_mean_meters, c(0.05, 0.95),na.rm = TRUE))+
  labs(title = "Quality score of coffee samples by altitude_mean_meters (Rescale)",
       x = "Quality Class",
       y = "altitude_mean_meters")
```

From the boxplot, we can analyse that as the mean altitude increases, the coffee quality also increases.




### Year Harvested
```{r plotting year harversted}
ggplot(data = coffee_data, aes(x = Qualityclass, group = harvested)) + 
  geom_bar(aes(y = ..prop.., fill = harvested), 
           stat = "count", position = "dodge") + 
  scale_color_gradient(low = "red", high = "green")+
  labs(x = "Quality of Coffee", y = "Proportion")
```


As can be seen above, the good quality coffee harvested more than poor quality coffee in total. But there is no significant difference between them.


In gengeral, we can find that there is no difference on means of Category 2 defects. Moreover, country of origin and year harvested are category variables. So the three variables can be removed to make the model more effective. But we need to investigate more to decide the final model.



## Formal Analysis {#sec:FA}

### Using Generalized Linear Model


Model 1: using all the available variables
```{r generalized model 1}

coffee_data$binary_response <- ifelse(coffee_data$Qualityclass == "Good", 1, 0)
mod.coffee1 <- glm(binary_response ~ country_of_origin +
                    aroma + flavor + acidity + category_two_defects +
                    altitude_mean_meters + harvested, 
                  data = coffee_data, family = binomial(link = "logit")) 

mod.coffee1 %>%
summary()
```


Firstly, we build a model with all variables.

Then, we exclude three suspected variables(country of origin, year harvested and category of two defect) one by one to build 3 models and compare them.


Model 2 (excluding **Country of origin**).
```{r generalized model 2}

mod.coffee2 <- glm(binary_response ~ aroma + flavor + acidity +  category_two_defects + altitude_mean_meters + harvested, 
                  data = coffee_data, family = binomial(link = "logit")) 
mod.coffee2 %>%
summary()
```


Model 3 (excluding **Country of origin** and **category two defect**).
```{r generalized model 3}

mod.coffee3 <- glm(binary_response ~ aroma + flavor + acidity +    altitude_mean_meters + harvested, 
                  data = coffee_data, family = binomial(link = "logit")) 
mod.coffee3 %>%
summary()
```


Model 4 (excluding **Country of origin**, **category two defect** and **year of harvested**)
```{r generalized model 4}

 
mod.coffee4 <- glm(binary_response ~ aroma + flavor + acidity + altitude_mean_meters, 
                  data = coffee_data, family = binomial(link = "logit")) 
mod.coffee4 %>%
summary()
```



### Building the model

From the distribution plot we learn that the distribution of the **country of origin** are relatively the same between **good and poor quality coffee**. We also learn that the mean of **Category two defect** is relatively the same as well for **good and poor quality coffee**. The same case also appears in the variable year harvested, we can see there is no significant difference between the distribution of **year harvested** in the **good and poor quality coffee**.

Hence we conduct 3 different model for this study, they are:

1. Model 1: Including all variable (i.e the country of origin, aroma, flavor,       acidity, the count of category two defect, the mean of altitude (in meters) and year harvested).

2. Model 2: Including all variable *except* the country of origin.

3. Model 3: Including all variable *except* the country of origin and the count of category two defect.

4. Model 4: Including all variable *except* the country of origin, the count of category two defect and the year harvested.

### Model Comparison 

From the models we try to compare the values in order to decide which model fits the data better. The summary of model is presented in the table below:
```{r comparing model}

model.comp.values.1 <- glance(mod.coffee1)
model.comp.values.2 <- glance(mod.coffee2)
model.comp.values.3 <- glance(mod.coffee3)
model.comp.values.4 <- glance(mod.coffee4)

Models <- c('Model 1','Model 2','Model 3', 'Model 4') 
bind_rows(model.comp.values.1, model.comp.values.2,
          model.comp.values.3,model.comp.values.4,.id="Model") %>%
  mutate(Model=Models) %>%  
  kable(
     digits = 2,
     caption = "Model comparison values for different models" 
  )
```

To choose the best fit for our model, we choose the model with low AIC, BIC and low deviance. From the table above, we can see that model 1 has the highest BIC, so model 3 is the best fit compared to the erst.



Now, lets look at the summary of model 3:
```{r summary model 3}

#Presenting summary model 3
summ(mod.coffee3)
```

Now, lets take a look at the confidence intervals of the explanatory variables:
```{r confidence interval}
confint(mod.coffee3)%>%
  kable()
```


We can see the p values for **altitude_mean_meters** and **harvested**  are 0.60 and 0.57 respectively, both are large than 0.05. The confident intervals for these
two variables also contain zero. That means, they are not significant. So, we drop the variables
**altitude_mean_meters** and **harvested**.


### Model summary
The final model contains three explanatory variables aroma, flavor and acidity.

We then check the summary of this **final model**:
```{r final model}
mod.final <- glm(binary_response ~ aroma + flavor + acidity, 
                  data = coffee_data, family = binomial(link = "logit")) 
mod.final %>%
summary()
```
We can see that all the variables are significant and the AIC value is 650.46


### Log odds plot

Finally, we look at the plot of the log odds of our final model:


```{r log odds plot}

#plotting log odds
plot_model(mod.final, show.values = TRUE, transform = NULL,
           title = "Log-Odds", show.p = FALSE)


```

We interpret the odds ratios as follows: 

- for each unit increase in the aroma grade of the coffee, their odds of being good quality coffee increases (by a factor of 4.30), 

- for each unit increase in the flavor grade of the coffee, the odds of it being good quality coffee increases (by a factor of 7.15), 

- the more acidic the coffee, their odds of being good quality coffee increases (by a factor of 3.16)

### Probability of the coffee being good quality by 3 factors:

```{r probability }
plot_model(mod.final, type = "pred", title = "Probability of coffee quality being good by the 3 different factors",
            axis.title = c(mod.final$colnumes,"Prob. of coffee quality being good")) 
```

As can be seen above, the coffee will become better if aroma, flavor and acidity become better. 
The three probability curves are similar, all rising steadily and then rising sharply around 6.








