---
title: "What is the factor of a good quality coffee"
author: "Group 12"
date: "2023-03-14"
#output: html_document
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA) 
```

```{r libraries, echo=FALSE}
library(tidyverse)
library(moderndive)
library(gapminder)
library(sjPlot)
library(stats)
library(jtools)
library(MASS)
library(knitr)
library(ggplot2)
library(dplyr)
library(janitor)
library(broom)
```

# Introduction {#sec:Intro}
For this research we want to know what is the factor that influence the quality of coffee. First we want to look at the distribution of the data.

```{r reading data,echo=TRUE}
coffee_data <- read.csv("dataset12.csv")
```


## Data distribution {#sec:EDA}

### The Country origin of the coffee

Below is the percentage of good and bad quality coffee for each origin of the coffee.
```{r check distribution origin}
#Check the percentage of good and poor quality coffee for each country of origin

coffee_data %>% 
  tabyl(country_of_origin, Qualityclass) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  adorn_ns()

```

From the figure below, we can see that the good quality coffee come from the countries with gradient red, orange and green, meanwhile the poor quality coffee come from the countries with gradient blue and pink.But the difference is not significant.

```{r plotting origin}

#Plot the distribution of coffee quality class based on the origin

ggplot(coffee_data, aes(y = Qualityclass)) +
  geom_bar(aes(fill = country_of_origin), position = position_stack(reverse = TRUE)) +
  theme(legend.position = "top")+
  scale_color_gradient(low = "yellow", high = "green")+
  labs(x = "Proportion of Country of Origin", y = "Quality of Coffee")
```








###  The Aroma of the coffee

Another variable of interest is the **aroma** of the coffee, here we want to see 
how the **aroma** of the coffee distributed around the quality of the coffee.


```{r plotting aroma1}

#Plot the distribution of coffee quality class based on the aroma
#First plot without rescaling
ggplot(data = coffee_data, aes(x = Qualityclass, y = aroma)) +
  geom_boxplot() +
  labs(title = "Quality score of coffee samples by aroma",
       x = "Aroma",
       y = "Qualityclass")
```


Rescale the distribution of Aroma of the coffee

It is hard to get the picture of the distribution, because of the present of outlier in this distribution. So we try to rescale the boxplot to get better visual of the distribution.

From the boxplot below we can see that poor quality coffee has lower value of the aroma of the coffee. 

Please note that for the rest of boxplot figure, we rescale them in order to make the clearer representation of the distribution.


```{r plotting aroma2}

#Rescaling the plot
ggplot(data = coffee_data, aes(x = Qualityclass, y = aroma)) +
  geom_boxplot() +
  scale_y_continuous(limits = quantile(coffee_data$aroma, c(0.05, 0.95)))+
  labs(title = "Quality score of coffee samples by aroma (Rescale)",
       x = "Quality Class",
       y = "Aroma")
```


### Flavour


```{r plotting flavor}
ggplot(data = coffee_data, aes(x = Qualityclass, y = flavor)) +
  geom_boxplot() +
  scale_y_continuous(limits = quantile(coffee_data$flavor, c(0.05, 0.95)))+
  labs(title = "Quality score of coffee samples by flavor (Rescale)",
       x = "Quality Class",
       y = "flavor")
```

As can be seen from the above boxplot, the poor quality coffee has lower flavour value than good quality coffee.



### Acidity


```{r plotting acidity}
ggplot(data = coffee_data, aes(x = Qualityclass, y = acidity)) +
  geom_boxplot() +
  scale_y_continuous(limits = quantile(coffee_data$acidity, c(0.05, 0.95)))+
  labs(title = "Quality score of coffee samples by acidity (Rescale)",
       x = "Quality Class",
       y = "acidity")
```

We can find that the acidity grade of poor quality coffee is much lower.


### category_two_defects


```{r plotting category two defect}
ggplot(data = coffee_data, aes(x = Qualityclass, y =    category_two_defects)) +
  geom_boxplot() +
  scale_y_continuous(limits =   quantile(coffee_data$category_two_defects, c(0.05, 0.95)))+
  labs(title = "Quality score of coffee samples by category_two_defects (Rescale)",
       x = "Quality Class",
       y = "category_two_defects")
```

The counts of category 2 type defects in the batch of coffee beans tested between good quality and poor quality coffee are similar. And both of them have many outliers.


### altitude_mean_meters 



```{r plotting altitude mean (meters)}
ggplot(data = coffee_data, aes(x = Qualityclass, y =    altitude_mean_meters)) +
  geom_boxplot() +
  scale_y_continuous(limits =   quantile(coffee_data$altitude_mean_meters, c(0.05, 0.95),na.rm = TRUE))+
  labs(title = "Quality score of coffee samples by altitude_mean_meters (Rescale)",
       x = "Quality Class",
       y = "altitude_mean_meters")
```

We can see that the mean altitude of the growers farm of good quality coffee is a little higher.



### Year Harvested
```{r plotting year harversted}
ggplot(data = coffee_data, aes(x = Qualityclass, group = harvested)) + 
  geom_bar(aes(y = ..prop.., fill = harvested), 
           stat = "count", position = "dodge") + 
  scale_color_gradient(low = "red", high = "green")+
  labs(x = "Quality of Coffee", y = "Proportion")
```

As can be seen above, the good quality coffee harvested more than poor quality coffee in total. But there is no significant difference between them.


In gengeral, we can find that the distributions of country of origin, year harvested and category of two defect are not significantly different. But we need to investigate more to decide the latest model.


## Formal Analysis {#sec:FA}

### Using Generalized Linear Model


+Model 1: using All variable
```{r generalized model 1}

coffee_data$binary_response <- ifelse(coffee_data$Qualityclass == "Good", 1, 0)
mod.coffee1 <- glm(binary_response ~ country_of_origin +
                    aroma + flavor + acidity + category_two_defects +
                    altitude_mean_meters + harvested, 
                  data = coffee_data, family = binomial(link = "logit")) 

mod.coffee1 %>%
summary()
```

Firstly, we build a model with all variables. However, we can try to remove insignificant variables to make the model more effective.

Then, we exclude three insignificant variables(country of origin, year harvested and category of two defect) one by one.



+Model 2 excluding **Country of origin**.
```{r generalized model 2}

mod.coffee2 <- glm(binary_response ~ aroma + flavor + acidity +  category_two_defects + altitude_mean_meters + harvested, 
                  data = coffee_data, family = binomial(link = "logit")) 
mod.coffee2 %>%
summary()
```


+Model 3 excluding **Country of origin** and **category two defect**.
```{r generalized model 3}

mod.coffee3 <- glm(binary_response ~ aroma + flavor + acidity +    altitude_mean_meters + harvested, 
                  data = coffee_data, family = binomial(link = "logit")) 
mod.coffee3 %>%
summary()
```


+Model 4 excluding **Country of origin**, **category two defect** and **year of harvested**.
```{r generalized model 4}

 
mod.coffee4 <- glm(binary_response ~ aroma + flavor + acidity + altitude_mean_meters, 
                  data = coffee_data, family = binomial(link = "logit")) 
mod.coffee4 %>%
summary()
```



### Building the model

From the distribution plot we learn that the distribution of the **country of origin** are relatively the same between **good and poor quality coffee**. We also learn that the mean of **Category two defect** is relatively the same as well for **good and poor quality coffee**. The same case also appears in the variable year harvested, we can see there is no significant difference between the distribution of **year harvested** in the **good and poor quality coffee**.

Hence we conduct 3 different model for this study, they are:

1. Model 1: Including all variable (i.e the country of origin, aroma, flavor,       acidity, the count of category two defect, the mean of altitude (in meters) and year harvested).

2. Model 2: Including all variable *except* the country of origin.

3. Model 3: Including all variable *except* the country of origin and the count of    category two defect.

4. Model 4: Including all variable *except* the country of origin, the count of    category two defect and the year harvested.

### Model Comparison 

From the models we try to compare the values in order to decide which model fit the data better. The summary of model is presented in the table below:
```{r comparing model}

model.comp.values.1 <- glance(mod.coffee1)
model.comp.values.2 <- glance(mod.coffee2)
model.comp.values.3 <- glance(mod.coffee3)
model.comp.values.4 <- glance(mod.coffee4)

Models <- c('Model 1','Model 2','Model 3', 'Model 4') 
bind_rows(model.comp.values.1, model.comp.values.2,
          model.comp.values.3,model.comp.values.4,.id="Model") %>%
  mutate(Model=Models) %>%  
  kable(
     digits = 2,
     caption = "Model comparison values for different models" 
  )
```

In order to choose the best fit for our model, we choose the model with low AIC, BIC and large deviance. From the table above, we can see that model 1 has the highest BIC, so the model 3 is the best fit according to the table .



After choosing model 3 from above, then we look at the summary of the model:
```{r summary model 3}

#Presenting summary model 3
mod.coffee3 %>% summary()
summ(mod.coffee3)
confint(mod.coffee3) %>%
  kable()
```


We can see the p values for **altitude_mean_meters** and **harvested** two 
variables are 0.60 and 0.57 respectively, both are large than 0.05. The CIs for this 
two variables are also contain the zero. According to this two reasons, we need drop
**altitude_mean_meters** and **harvested**.


### Model summary
The final model contains aroma flavor and acidity this three explanatory variables.

we need summary the **final model** again:
```{r final model}
mod.final <- glm(binary_response ~ aroma + flavor + acidity, 
                  data = coffee_data, family = binomial(link = "logit")) 
mod.final %>%
summary()
```


### Log odds plot

Finally we look at the plot of the log odds from our model


```{r log odds plot}

#plotting log odds
plot_model(mod.final, show.values = TRUE, transform = NULL,
           title = "Log-Odds", show.p = FALSE)


```

We interpret the odds ratios as follows: 

- for each increase in the aroma value of the coffee, their odds of being good quality   coffee increase (by a factor of 4.30), 

- for each increase in the flavor value of the coffee, their odds of being good         quality coffee increase (by a factor of 7,15), 

- the more acidic the coffee, their odds of being good quality coffee increase (by a    factor of 3.16), 


```{r probability }
plot_model(mod.final, type = "pred", title = "Probability of coffee quality being good by three factors",
            axis.title = c(mod.final$colnumes,"Prob. of coffee quality being good")) 
```


As can be seen above, the coffee will become better if aroma, flavor and acidity become better. 
The three probability curves are similar, all rising steadily and then rising sharply around 6.







